Using evolutionary algorithms to optimize the estimation of a diffusion process\titlenote{(Produces the permission block, and copyright information). For use with SIG-ALTERNATE.CLS. Supported by ACM.}

### 2015-02-06 paperGecco.tex
Cambios en la introduccion, faltan las formulas

\documentclass{sig-alternate}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx, epsfig}

\usepackage{url}
\urldef{\mailsa}\path|{mgarenas,nrico,deromero,pacv,jmerelo}@ugr.es|

\begin{document}
\conferenceinfo{GECCO'15,} {July 11-15, 2015, Madrid, Spain.}
    \CopyrightYear{2015}
    \crdata{TBA}
    \clubpenalty=10000
    \widowpenalty = 10000
%\mainmatter  % start of an individual contribution
%

\title{Using evolutionary algorithms to optimize the estimation of a diffusion process}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{6} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Nuria Rico\titlenote{Dr.~Nuria Rico Castro.}\\
       \affaddr{Faculty of Science University of Granada}\\
       \email{nrico@ugr.es}
% 2nd. author
\alignauthor
M.G. Arenas\titlenote{Dr. ~ M. Isabel García Arenas.}\\
       \affaddr{ETSIIT - University of Granada}\\
       \email{mgarenas@ugr.es}
% 3rd. author
\alignauthor
D. Romero\titlenote{Dr. ~Desirée Romero.}\\
       \affaddr{ETSIIT - University of Granada}\\
       \email{deromero@ugr.es}
\alignauthor
P. A. Castillo\titlenote{Dr. ~Pedro A. Castillo Valdivieso.}\\
       \affaddr{ETSIIT - University of Granada}\\
       \email{pacv@ugr.es}
\alignauthor
J. M. Crespo\titlenote{Dr. ~J. M. Crespo.}\\
       \affaddr{ETSIIT - University of Granada}\\
       \email{crespo@ugr.es}
%\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor J. J. Merelo\\
       \affaddr{Dr. Juan Julián Merelo Guervós}\\
       \affaddr{ETSIIT-University of Granada}\\
       \email{jjmerelo@ugr.es}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

Many probabilistic models are frequently used for natural growth-pattern modeling and making forecast. Although there exists a lot of different growth-patterns, and many models to use, among the widely studied and useful models are diffusion processes. They are frequently employed in biological growth modeling \cite{lognormal}.

In this paper we focus our attention on the Gompertz-lognormal diffusion process \cite{proceedingBol}. It can be seen as a mixture between a Gompertz-type process \cite{Gut07b} and a lognormal process \cite{GutCy} since the infinitesimal moments that define the Gompertz-lognormal diffusion process are:

%{\scriptsize
\[
\begin{array}{l}
A_1(x,t)=(m e^{-\beta t}+c)x \\
A_2(x,t)=\sigma^2 x^2,
\end{array}
\]
with $m > \beta > 0$, $c\in \mathbb{R}$.
%}
%Maribel: Deberíamos poner qué es m, que es e, que es c y por supuesto mencionar que t es el tiempo y que x es la variable independiente no?

@@

To carry out the estimation, we consider the transition density function, that depends on $m$, $\beta$, $c$ and $\sigma^2$ parameters. Estimating the set of parameters leads to the estimation of the main features of the process. The most widely used, specifically for predicting purposes, are the mean, mode and quantile functions. Their expressions depend on the set of parameters $\{m;\beta; c; \sigma^2\}$ and, in the case of a lognormal initial distribution being considered, on the $\{ \mu_1, \sigma_1^2 \}$ set too.

@@

\bibitem{proceedingBol}
Guti{\'e}rrez, R., Rico, N., Rom{\'a}n, P., Romero, D., Torres, F.:
Un nuevo proceso de difusi{\'o}n para modelizar patrones de crecimiento mixto exponencial-Gompertz. XXX Congreso Nacional SEIO. Libro de Actas, pp. 1--15 (2007)

\bibitem{Gut07b}
Guti{\'e}rrez, R., Rom{\'a}n, P., Romero, D., Serrano, J.J. y Torres, F.:
A new Gompertz-type diffusion process with application to random growth. Math. Biosci. 208,147--165 (2007)

\bibitem{GutCy}
Guti{\'e}rrez, R., Rom{\'a}n, P., Romero, D. and Torres, F.:
Applications of the univariate lognormal diffusion process with exogenous factors in forecasting. {Cybernet Syst.} {34}(8),709--724 (2003)

\bibitem{Gut99} Guti{\'e}rrez, R., Rom{\'a}n, P. and Torres, F.:
Inference and first-passage-times for the lognormal diffusion process with exogenous factors: application to modelling
in economics. Appl. Soch. Model. Bus. {15},325-332 (1999)

\bibitem{SA}
Ingber, L.:
Simulated annealing: Practice versus theory. Mathl. Comput. Modelling, vol. 18(11), pp. 29--57 (1993)

\bibitem{Kirkpatrick83}
S. Kirkpatrick1, C. D. Gelatt Jr. M. P. Vecch:
Optimization by Simulated Annealing. Science Vol. 220 no. 4598 pp. 671-680 (1983)

\bibitem{Michalewicz96}
Zbigniew Michalewicz:
Heuristic methods for evolutionary computation techniques. J. Heuristics 1(2):177-206 (1996)

@@

\bibitem{Kloeden}
Kloeden, P. E., Platen, E. y Schurz, H.:
Numerical solution of SDE through computer experiments, {Springer-Verlag} (1994)

\bibitem{Rao}
Rao, N. J., Borwankar, J. D. y Ramkrishna, D.:
Numerical solution of Ito integral equations. {SIAM Journal on Control and Optimization} {12}, 124--139 (1974)

\bibitem{error}
De Gooijer, J.G., Hyndman, R.J.:
25 Years of IIF Time Series Forecasting: A Selective Review. Tinbergen Institute Discussion Papers No. TI 05-068/4 . (2005)

\bibitem{GProp}
P. A. Castillo, J. Carpio, J. J. Merelo, V. Rivas, G. Romero, and A. Prieto:
Evolving Multilayer Perceptrons, Neural Processing Letters, vol. 12, no. 2, pp. 115--127, 2000.

\bibitem{GPropII}
S. Fahlman:
Faster-learning variations on back-propagation: An empirical study, in Proceedings of the 1988 Connection- ist models summer school. M. Kaufmann, 1988, pp. 38--51.

\bibitem{lognormal}
Log-normal Distributions across the Sciences: Keys and Clues
Article in Bioscience 51, No. 5, May 2001, 341-352.

\bibitem{blx}
Manuel Lozano, Francisco Herrera, Natalio Krasnogor, and Daniel Molina:
Real-coded memetic algorithms with crossover hill-climbing. Evol. Comput. 12, 3. Sep, 2004, 273-302.

\bibitem{VNS} Mladenovic, N. and Hansen, P.: Variable neighborhood search. {Computers and Operations Research} {24}, 1097--1100 (1997)

\end{thebibliography}

\end{document}

>>>>

\documentclass{sig-alternate}
\usepackage{multicol}
\usepackage{graphicx, epsfig}
\usepackage{graphicx}
\begin{document}
\conferenceinfo{GECCO'15,} {July 11-15, 2015, Madrid, Spain.}
    \CopyrightYear{2015}
    \crdata{TBA}
    \clubpenalty=10000
    \widowpenalty = 10000
%
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Using evolutionary algorithms to optimize the estimation of a diffusion process\titlenote{(Produces the permission block, and
copyright information). For use with
SIG-ALTERNATE.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{6} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
%Nuria Rico\titlenote{Dr.~Nuria Rico Castro.}\\
%       \affaddr{Faculty of Science }\\
%       \affaddr{University of Granada}\\
%       \email{nrico@ugr.es}       
Nuria Rico\titlenote{Dr.~Rico}\\
       \affaddr{Statistics and Operational Research}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{nrico@ugr.es}
% 2nd. author
\alignauthor
M. G. Arenas\titlenote{Dr. ~Arenas}\\
       \affaddr{Computer Architecture}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{mgarenas@ugr.es}
% 3rd. author
\alignauthor 
Desir\'{e}e Romero\titlenote{Dr. ~Romero}\\
       \affaddr{Statistics and Operational Research}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{deromero@ugr.es}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor 
P. A. Castillo\titlenote{Dr. ~Castillo}\\
       \affaddr{Computer Architecture}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{pacv@ugr.es}
% 5th. author
\alignauthor
J. M. Crespo\titlenote{Dr. ~Crespo}\\
       \affaddr{Statistics and Operational Research}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{crespo@ugr.es}
% 6th. author
\alignauthor
J. J. Merelo\titlenote{Dr. ~Merelo}\\
       \affaddr{Computer Arquitecture}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{jmerelo@ugr.es}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{1 Feb 2015}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

Many probabilistic models are frequently used for natural growth-pattern modeling and making forecast. Although there exists a lot of different growth-patterns, and many models to use, among the widely studied and useful models are diffusion processes. They are frequently employed in biological growth modeling. \cite{lognormal}.

%%%Nuevo

Diffusion process is one usual mathematical formulation of a model for population growth in a randomly
varying environment \cite{lognormal}. These models describe how a population adopts a new innovation, technology or behaviour. With these kind of models, we have the transity density function (t.d.f.) which express the chance of observing
any size for the population at any time when the initial size is specified. With the t.d.f we also are able to determine the time at that population first reaches a fixed size. Moreover, we can forecast future values of the population size as well as confidence intervals for the predictions.

In this paper we focus our attention on the Gompertz-lognormal diffusion process \cite{proceedingBol}. It can be seen as a mixture between a Gompertz-type process \cite{Gut07b} and a lognormal process \cite{GutCy}. The first one models precesses with trends showing enclosed growth patterns, as biological growth with finite space or resources. The second one is useful for modelling not bounded growth with exponential trend.

The mixture of them, the Gompertz-lognormal diffusion process, combines both trends; a first part Gompertz-like growth and a second exponential part. This process is defined through the infinitesimal moments
%{\scriptsize
\[
\begin{array}{l}
A_1(x,t)=(m e^{-\beta t}+c)x \\
A_2(x,t)=\sigma^2 x^2,
\end{array}
\]
with $m > \beta > 0$, $c\in \mathbb{R}$.
%}
%Maribel: Deber{\'\i}amos poner qu{\'e} es m, que es e, que es c y por supuesto mencionar que t es el tiempo y que x es la variable independiente no?

@@

To carry out the estimation, we consider the t.d.f., that depends on $m$, $\beta$, $c$ and $\sigma^2$ parameters. Estimating the set of parameters leads to the estimation of the main features of the process. The most widely used, specifically for predicting purposes, are the mean, mode and quantile functions. Their expressions depend on the set of parameters $\{m;\beta; c; \sigma^2\}$ and, in the case of a lognormal initial distribution being considered, on the $\{ \mu_1, \sigma_1^2 \}$ set too.

@@

\bibitem{proceedingBol}
Guti{\'e}rrez, R., Rico, N., Rom{\'a}n, P., Romero, D., Torres, F.:
Un nuevo proceso de difusi{\'o}n para modelizar patrones de crecimiento mixto exponencial-Gompertz. XXX Congreso Nacional SEIO. Proceedings book, pp. 1--15 (2007)

\bibitem{Gut07b}
Guti{\'e}rrez, R., Rom{\'a}n, P., Romero, D., Serrano, J.J.,  Torres, F.:
A new Gompertz-type diffusion process with application to random growth. Math. Biosci. 208, 147--165 (2007)

\bibitem{GutCy}
Guti{\'e}rrez, R., Rom{\'a}n, P., Romero, D., Torres, F.:
Applications of the univariate lognormal diffusion process with exogenous factors in forecasting. {Cybernet Syst.} {34}(8), 709--724 (2003)

\bibitem{Gut99} Guti{\'e}rrez, R., Rom{\'a}n, P., Torres, F.:
Inference and first-passage-times for the lognormal diffusion process with exogenous factors: application to modelling
in economics. Appl. Soch. Model. Bus. {15}, 325-332 (1999)

\bibitem{SA}
Ingber, L.: 
Simulated annealing: Practice versus theory. Mathl. Comput. Modelling, 18(11), pp. 29--57 (1993)

\bibitem{Kirkpatrick83}
S. Kirkpatrick1, C. D. Gelatt Jr. M. P. Vecch:
Optimization by Simulated Annealing. Science 220(4598) pp. 671--680 (1983)

\bibitem{Michalewicz96}
Michalewicz, Z.:
Heuristic methods for evolutionary computation techniques. J. Heuristics 1(2) pp. 177--206 (1996)

@@

\bibitem{Kloeden}
Kloeden, P. E., Platen, E., Schurz, H.:
Numerical solution of SDE through computer experiments, {Springer-Verlag} (1994)

\bibitem{Rao}
Rao, N. J., Borwankar, J. D., Ramkrishna, D.:
Numerical solution of Ito integral equations. {SIAM Journal on Control and Optimization} {12}, 124--139 (1974)

\bibitem{error}
De Gooijer, J.G., Hyndman, R.J.:
25 Years of IIF Time Series Forecasting: A Selective Review. Tinbergen Institute Discussion Papers No. TI 05-068/4  (2005)

\bibitem{GProp}
Castillo, P. A., Carpio, J., Merelo, J., Rivas, V., Romero, G., Prieto, A.: 
Evolving Multilayer Perceptrons, Neural Processing Letters, 12(2), pp. 115--127, (2000)

\bibitem{GPropII}
Fahlman, S.: 
Faster-learning variations on back-propagation: An empirical study, in Proceedings of the 1988 Connection- ist models summer school. M. Kaufmann pp. 38--51 (1988)

\bibitem{lognormal}
Limpert, E., Stahel, WA., Abbt, M.: 
Log-normal Distributions across the Sciences: Keys and Clues, Bioscience 51(5) pp. 341--352 (2001)

\bibitem{blx}
Lozano, M., Herrera, F., Krasnogor, N., Molina D.:
Real-coded memetic algorithms with crossover hill-climbing. Evol. Comput. 12(3) pp. 273--302 (2004)

\bibitem{VNS} 
Mladenovic, N., Hansen, P.: 
Variable neighborhood search. {Computers and Operations Research} {24}, 1097--1100 (1997)

\end{thebibliography}
\end{document}
